Uncertainty in social learning evolution

### 2022-05-23 ms.tex
complete model and analysis; stubbed rest with new figs and old tables

Social learning is essential to survival. It is likely to evolve when it is more
efficient than asocial, trial-and-error learning. Theoretically, some uncertainty is
necessary for social learning to be necessary, but too much uncertainty makes
social information useless. This fact is empirically
supported across biology and human sciences. However, we lack a theoretical
framework to predict the effects of specific classes and types of uncertainty on social
learning evolution. Furthermore, existing models and experimental operationalizations of
uncertainty are ambiguously related, and 
tend to only consider a small number of sources of uncertainty and behavioral
choices.  Here we use evolutionary agent-based modeling to consolidate
models of uncertainty in social learning evolution to improve on these
shortcomings. We model a time-varying environment with a varying number
of possible behaviors agents could perform to acquire payoffs.
We show that environmental variability, ambiguous payoffs, 
larger decision sets, and shorter agent lifespans interact to 
affect social learning evolution in complex ways that 
sometimes caused evolution to be highly uncertain about whether social learning was
the optimal strategy. 
Our work strengthens social evolution theory that could help 
guide human, non-human, and artificial intelligence towards optimal responses 
to existential threats and new opportunities.\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning is essential to human and other species' everyday life and survival.
It allows individuals to solve problems when acquiring information from others is
more efficient than learning on one's own~\cite{Laland2004}. Theory predicts that
social learning should be favored in contexts with greater
uncertainty~\cite{BoydRicherson1985,Henrich1998}, up to a
point~\cite{Rogers1988,Feldman1996}; this prediction has received some empirical
support across species~\cite{McElreath2005,Kendal2018,Allen2019}.  However, the
meaning of the term \emph{uncertainty} often conflates environmental variability,
spatial heterogeneity, ambiguity or uncertainty about payoff structure, and other
variability that could be plausibly interpreted as \emph{uncertainty}.  Many models
of social learning evolution do not account for 
individual-level cognition~\cite{Heyes2016}, though humans
clearly have evolved cognitive mechanisms for dealing with 
uncertainty generally~\cite{Gershman2019,Schulz2019}.
Furthermore, social learning evolution theory tends to be organized around an
increasing number of modeling choices~\cite[Figure 1]{Kendal2018} instead
of around broader contextual assumptions such as uncertainty conditions. It remains
an outstanding question, then, to identify how different contexts and forms of
uncertainty might combine to affect social learning evolution in complex ways.
Answering this question is critical for understanding human behavior adapting to a
rapidly changing world, both to mitigate existential
threats~\cite{Moya2020,Jones2021}, and to captialize on new opportunities such as
transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.
We will answer this question about how uncertainty drives evolution, including
identifying regimes where evolution itself is highly uncertain and often 
leads populations into suboptimal learning strategies~\cite{Rogers1988}.

\mt{This paragraph is a first try to introduce the four uncertainty dimensions
with examples to illustrate the problem we are answering.
I outlined the next paragraph to get into more details of social learning:
how cultural evolution works; conformity, success-biased, etc. Cristina, please
edit this.} 
To gain a more systematic understanding of uncertainty and social learning
evoluiton, we identified four common classes of ecological
uncertainty (``principal components'' of uncertainty) used as factors
in social learning evolution models: (1) environmental variability; (2) payoff
ambiguity (difference between best behavior choice and others); (3) 
selection set size (number of ecologically possible behaviors); and 
(4) the lifespan of agents. These were chosen because modeling and empirical 
study designs across taxa tend use one or more of these uncertainty dimensions.
Environmental variability is sometimes modeled as catastrophic and 
happens only once in an individual's lifetime~\cite{Rogers1988}, or may be
modeled as occurring with a fixed or variable
probability~\cite{Feldman1996,McElreath2005}---in these examples environments
can only be in one of two states. Payoff ambiguity is sometimes in the form
of differences in expected payoffs arising from different
behaviors or strategies~\cite{Enquist2007,Rendell2010}
or by varying the standard deviation of payoffs~\cite{McElreath2005}. 
Selection set size is often limited to one size throughout social learning
evolution studies: empirical studies of \emph{bombus terresteris} (bumble
bee)~\cite{Baracchi2018} and \emph{parus major} (great tit)~\cite{Aplin2017}
populations each provided two behaviors for
the bees and birds to perform, each yielding greater or smaller
payoffs depending on experimental treatment; studies in humans have used two
or three possible behaviors~\cite{McElreath2005,Toyokawa2019}; 
modeling studies have used two 
behaviors~\cite{Feldman1996,Rendell2010} and four behaviors~\cite{Enquist2007} 
in understanding the role of uncertainty. The age or lifespan of individuals was
found to affect on social learning in both \emph{parus
major}~\cite{Aplin2017} and \emph{poecilia reticulata} 
(guppies)~\cite{Leris2016} \mt{Need better connection between lifespan and 
age, or better references to support use of lifespan}.
Shorter lifespans mean individuals
die more uncertain of which behaviors are optimal, and pass that uncertainty on
to their offspring. To answer our research questions about the combined effect
of various forms of uncertainty on social learning evolution, we developed an
agent-based model that incorporates all four of
these uncertainty variables. We systematically vary these parameters
to understand and predict their effects on social learning evolution.

@@

\paragraph{Meaning of uncertainty in ecology and cognitive and social sciences} (JAMIE)
We need to support our claim that these four classes of uncertainty really 
count as forms of uncertainty and show how it is consistent across ecology and
cognitive and social science.

\paragraph{ABMs for social learning evolution theory development} (PAUL)

@@

To answer our research questions
we developed an agent-based model of social learning evolution in which a society of
$N$ individuals each must decide which of $B$ behaviors to perform at each time step
within a generation of $L$ time steps. At the end of each generation, agents are selected
with replacement from the population to reproduce another $N$ agents. Agents who
inherit the social learning trait then learn about behavioral payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with probabilistic
payoffs~\cite{SuttonBartoBook,McElreath2005,Rendell2010,Schulz2019}.  Each behavior
$b$ yields Bernoulli payoffs: payoff of 1 with probability $\pi_b$ and zero payoff
with probability $1 - \pi_b$; $\pi_b$ is therefore also the expected payoff of
behavior $b$. One behavior in each generation is optimal and pays off $\pihigh$,
while the rest pay off $\pilow < \pihigh$.  Model agents must decide which behavior to
perform at each time step.  To do this, agents use an explore-exploit strategy to
sometimes try the most profitable behavior they know about, and other times try
alternatives that may pay off more reliably.  We developed a series of computational
analyses by systematically varying uncertainty parameters and observing the
frequency that model populations evolve to be social learners, the average payoffs
accrued by agents in each setting, and the time it takes for model populations to
reach fixation (agents all social or all asocial learners).

@@

To guide behavior selection, agent $i$ counts how many times it
performed each behavior $b$, denoted $c_{ib}$. Agents use this count to 
update the mean payoff they have obtained from behavior $b$, denoted $\ledger$.
$\ledger$ is initialized to 0 for all $b$ at model initialization for
all agents and after each generation for individual learner agents. Social
learning agents' $\ledger$ is initialized as that of its teacher selected
through performance-biased oblique learning. Agent $i$'s accumulated payoffs
from performing several behaviors over their lifetime of $L$ time steps is
denoted $\pi_{i}$. If an agent $i$ is a social learner, then we write $s_i = 1$,
otherwise the agent is an asocial learner and $s_i = 0$.

At the beginning of each simulation model dynamics begin with 
one of the behaviors selected to 
yield expected payoff $\pihigh$, with the other $B-1$ behavior yielding
expected payoffs $\pilow$. At each time step, agents perform behavior $b$ 
with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}},
\end{equation}
\noindent
where $\beta$ is a parameter that adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$). 
Agents probabilistically recieve a payoff of 0 or 1 depending on whether the
bandit paid off, equivalent to a draw from a Bernoulli distribution with 
mean $\pi_b$. For concreteness we denote this probabilistic payoff
$\bandit$. On performing behavior $b$, agent $i$ updates the
corresponding behavior count by 1, $c_{ib} \leftarrow c_{ib} + 1$, and then
the expected payoffs calculated for that behavior are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}}.
\end{equation}
\noindent
This procedure continues for $L$ time steps within each generation.

In between generations agents first reproduce via asexual haploid reproduction; 
then new child agents learn from the
previous generation if they are social learners; and finally all agents from the
previous generation then die off. $N$ reproducers are selected with replacement, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
Child agents inherit their parents social learning trait $s_i$ without mutation.
Child agents with $s_i = 1$ then learn from one teacher from their parent's
generation, including possibly their parent. Child social learners select
their teacher first by selecting $N_T$ prospective teachers from the 
population completely at random, then selecting the one with highest
net payoffs $\pi_i$ for as their teacher, with ties broken randomly 
($N_T = 5$ in the main text;
\mt{Other values will be tested in the supplement}). While it is somewhat
arbitrary to first subset $N_T$, this is a conservative choice that
represnts the fact that access to the entire population is not generally
guaranteed. Social learners adopt their chosen teacher's ledger of 
expected payoffs $\ledger$, and set
all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$. 
Asocial learners ($s_i = 0$) are initialized with all $c_{ib} = 0$ and
$\ledger = 0$. The newly initialized population then begins behavior selection
and payoff accumulation for another $L$ time steps and the evolutionary process
continues until $\sum_i s_i = 1$ or $\sum_i s_i = 0$ \mt{I believe
all trials will eventually end like this, so in the next set of model runs I will
not allow early termination before one of these conditions are met}.

To analyze the effect of the four principal uncertainty factors, we systematically
varied their values and observed three key outcome variables aggregated across 1000
trial simulations, and across all agents. We varied $u \in \{0.0,
0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L
\in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We observed
three outcome measures: (1) $\meansl$, the average value of $s_i$ over all agents
and trials; (2) $\meanpi / L$, the average net payoffs within agent lifetimes over
trials, normalized by their lifespan; and (3) $\meanT / L$, or the mean number
of time steps to fixation normalized by lifespan\footnote{Equivalent to the mean
number of generations to fixation} 
(fixation means $\sum_i s_i = 0 \text{ or } 1$). \mt{Make sure to qualitatively
introduce these and their evolutionary importance in the intro.}
We inspect the outcomes through visualization environmental
variation is on the x-axis since it is the best understood: we expect that as $u$
increases, social learning prevalence $\meansl \to 0$. We expected and observed
that the other three uncertainty factors shift at which $u$ $\meansl$ starts to
decrease, and how rapidly $\meansl$ decreases as $u$ increases. Note the
hypothesis-testing concept of significance is meaningless here because we could
make any small outcome difference significant by running more simulation trials.
The important outcomes for this study are the patterns of our outcome 
variables over the systematic variation of the uncertainty factors.

@@

@@

\paragraph{Analysis 2: evolutionary uncertainty when $\meansoc \approx \meanasoc$}

\begin{figure}
  \caption{Mean payoff normalized for lifespan (y-axes, solid line with circles)
    generally track either expected all-social-learners payoff ($\meansoc$, dash-dot
    line with diamonds) or expected all-asocial payoff ($\meanasoc$, long dash
    horizontal lines) as $u$ increases (x-axes). When $\meansl$ goes from 1 to
  0 is when social learning payoffs approach individual learning payoffs due to
increased environmental variability.} 
  \label{fig:netPayoffs}
\centering
    \includegraphics[width=0.75\textwidth]{Figures/meanNetPayoffs.pdf}
\end{figure}

@@

By identifying four common uncertainty dimensions we were able to develop a model
that predicts how they might interact to foster or suppress social learning.
The model reproduced classic predictions that social learning can yield higher
payoffs than individual learning, but only up to some upper uncertainty bound.
By disambiguating and organizing various operationalizations of \emph{uncertainty},
we were able to add more detail to our theoretical understanding of which forms
of uncertainty interact in which ways to make social learning more or less
beneficial compared to asocial learning. When expected social learning payoffs
approached asocial learning payoffs the system exhibited far stronger drift; when
social or asocial learning were more obviously better the systems more consistently
evolved to the more profitable state. Note that in many cases we made modeling choices
that excluded some important mechanisms of social learning; for example we 
assumed success-biased learning, although conformity is theoretically important
as well~\cite{Muthukrishna2016a,Smaldino2018b}. Nonetheless, we believe our model
advances theoretical parsimony; and the software implementation is designed to
facilitate testing alternate modeling choices.

\paragraph{Disc. 2: Foundation for studying social learning evolution with group
structure (PAUL, JAMIE, CRISTINA)}~\cite{Katsnelson2014}.

>>>>

Social learning is essential to survival. It is likely to evolve when it is more
efficient than asocial, trial-and-error learning. Theoretically, social learning
is adaptive for some uncertainty, but too much uncertainty makes
social information unreliable. This fact is empirically
supported across biology and human sciences. However, it is unclear how
specific classes and types of uncertainty affect social
learning evolution. Furthermore, existing models and experimental operationalizations of
uncertainty are ambiguously related, and 
tend to only consider a small number of behavioral
choices.  Here we use evolutionary agent-based modeling to consolidate
models of uncertainty in social learning evolution to address these issues.
We model societies of agents who perform one of a number
of possible behaviors to acquire payoffs in a time-varying environment.
We identified complex patterns in social learning evolution depending on contextual
uncertainty variables: environmental variability, ambiguous payoffs, 
decision sets size, and effective lifespans. 
Our work advances social learning evolution theory in a way that could help 
guide human, non-human, and artificial intelligence towards optimal responses 
to existential threats and new opportunities.\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning is essential to human and other species' everyday life and survival.
Social learning enhances problem solving when acquiring information from others is
more efficient than learning on one's own~\cite{Laland2004}. Theory predicts that
social learning should be favored in uncertain
environments~\cite{BoydRicherson1985,Henrich1998}, up to a
point~\cite{Rogers1988,Feldman1996}; this prediction has received some empirical
support across species~\cite{McElreath2005,Kendal2018,Allen2019}.  However, the
meaning of the term \emph{uncertainty} often conflates environmental variability,
spatial heterogeneity, ambiguity in behavioral payoff structure, and other
variability that could plausibly be interpreted as \emph{uncertainty}.  Furthermore,
many models of social learning evolution do not account for individual-level
cognition~\cite{Heyes2016}, though humans clearly have evolved cognitive mechanisms
for dealing with uncertainty generally~\cite{Vul2014,Gershman2019,Schulz2019}.
Social learning evolution theory tends to be organized around an increasing number
of modeling choices~\cite[Figure 1]{Kendal2018} instead of around questions of when
and why social learning is adaptive, in which contexts.  Therefore, we developed
an evolutionary, empirically-motivated, agent-based model that operationalized
four principal uncertainty components to generate predictions of when 
social learning would evolve in which circumstances, and why. Answering this
question is critical for understanding human behavior adapting to a rapidly changing
world, both to mitigate existential threats~\cite{Moya2020,Jones2021}, and to
captialize on new opportunities such as transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

\mt{This paragraph is a first try to introduce the four uncertainty dimensions
with examples to illustrate the problem we are answering.
I outlined the next paragraph to get into more details of social learning:
how cultural evolution works; conformity, success-biased, etc. Cristina, please
edit this.} 
To gain a more systematic understanding of uncertainty and social learning
evoluiton, we identified four common classes of ecological
uncertainty (``principal components'' of uncertainty) used as factors
in social learning evolution models: (1) environmental variability; (2) payoff
ambiguity (difference between best behavior choice and others); (3) 
selection set size (number of ecologically possible behaviors); and 
(4) the lifespan of agents. These were chosen because modeling and empirical 
study designs across taxa tend use one or more of these uncertainty dimensions.
Environmental variability is sometimes modeled as catastrophic and 
happens only once in an individual's lifetime~\cite{Rogers1988}, or may be
modeled as occurring with a fixed or variable
probability~\cite{Feldman1996,McElreath2005}---in these examples environments
can only be in one of two states. Payoff ambiguity sometimes takes the form
of differences in expected payoffs arising from different
behaviors or strategies~\cite{Enquist2007,Rendell2010}
or by varying the standard deviation of payoffs~\cite{McElreath2005}. 
Selection set size often takes only one value: 
for example, in empirical studies of \emph{bombus terresteris} (bumble
bee)~\cite{Baracchi2018} and \emph{parus major} (great tit)~\cite{Aplin2017},
experimenters provided two behaviors for
the bees or birds to perform, respectively, each yielding greater or smaller
payoffs depending on experimental treatment. Human studies have used two
or three possible behaviors~\cite{McElreath2005,Toyokawa2019}.
Modeling studies have used two 
behaviors~\cite{Feldman1996,Rendell2010} and four behaviors~\cite{Enquist2007}. 
The age or lifespan of individuals was
found to affect on social learning in both \emph{parus
major}~\cite{Aplin2017} and \emph{poecilia reticulata} 
(guppies)~\cite{Leris2016} \mt{Need better connection between lifespan and 
age, or better references to support use of lifespan}.
Shorter lifespans mean individuals
die more uncertain of which behaviors are optimal, and pass that uncertainty on
to their offspring. To answer our research questions about the combined effect
of various forms of uncertainty on social learning evolution, we developed an
agent-based model that incorporates all four of
these uncertainty variables. We systematically vary these parameters
to understand and predict their effects on social learning evolution.

@@

\paragraph{Meaning of uncertainty in ecology and cognitive and social sciences } (JAMIE)
We need to support our claim that these four classes of uncertainty really 
count as forms of uncertainty and show how it is consistent across ecology,
and maybe connect to uncertainty in cognitive and social sciences more broadly.

\paragraph{Evolutionary ABMs for social learning evolution theory development} (PAUL)

@@

\begin{table}[h]
    \caption{Dynamic agent-level variables. Each has an implicit time dependence.}
    \label{tab:modelParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

        Attribute & Description & Initial value \\

        \midrule

        $s_i$  & Social learner trait: 1 if agent $i$ is social learner; 0 otherwise & 0
        or 1 equally likely \\

        $\bar\pi_{ib}$ & ``Ledger'': mean payoffs acquired via behavior $b$ by $i$ 
                       & $B$-vector of zeros \\

        $c_{ib}$ & Count of how many times agent $i$ performed $b$ 
              & $B$-vector of zeros \\

        $\pi_i$ & Net payoffs accumulated by $i$ within generation
                                & 0.0 \\
        \bottomrule
    \end{tabular}
\end{table}

To answer our research questions we developed an agent-based model of social
learning evolution in which a society of $N$ individuals each must decide which of
$B$ behaviors to perform at each time step within a generation of $L$ time steps. At
the end of each generation, agents are selected with replacement from the population
to reproduce another $N$ agents. Agents who inherit the social learning trait then
learn about behavioral payoffs from the previous generation, while asocial learners
begin life with no \emph{a priori} knowledge of the world.  Each behavior is a
``bandit'', a common modeling and experimental approach for representing behaviors
with probabilistic
payoffs~\cite{SuttonBartoBook,McElreath2005,Rendell2010,Schulz2019}.  Each behavior
$b$ yields Bernoulli payoffs: payoff of 1 with probability $\pi_b$ and zero payoff
with probability $1 - \pi_b$; $\pi_b$ is therefore also the expected payoff of
behavior $b$. One behavior in each generation is optimal and pays off $\pihigh$,
while the rest pay off $\pilow < \pihigh$.  Model agents must decide which behavior
to perform at each time step.  To do this, agents use an explore-exploit strategy to
sometimes try the most profitable behavior they know about, and other times try
alternatives that may pay off more reliably.  We developed a series of computational
analyses by systematically varying uncertainty parameters and observing the
frequency that model populations evolve to be social learners, the average payoffs
accrued by agents in each setting, and the time it takes for model populations to
reach fixation (agents all social or all asocial learners). Model dynamics are
illustrated in Figure~\ref{fig:schematic}.

\begin{figure}
  \caption{Agents are initialized as randomly social learners or not, with their
  payoff observations all initialized to zero (A). Then agents begin selecting
and performing behaviors and accumulating payoffs, which goes on for $L$
timesteps (B). After $L$ time steps, agents are selected to reproduce,
social learner children learn from a member of their parent's generation, and
the previous generation dies off (C). The simulation stops if children are all
social or asocial learners (i.e.\ the system reaches fixation), 
otherwise repeat another generation and evolution (3).}
  \label{fig:schematic}
  \centering
    \includegraphics[width=\textwidth]{Figures/IntraInterGenerationalDynamics.pdf}
\end{figure}

@@

\vspace{2em}
\begin{table}[h]
    \caption{Global environmental and cognitive initialization parameters.
      Bold indicates \emph{default value tested}.
    \mt{TODO: Test more $N_\text{T}$ for sensitivity analysis}}
    \label{tab:modelParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

        Symbol & Description & Values tested \\

        \midrule

        $B$       & Number of possible behaviors (represented by ``bandits'') 
                  & 2, 4, 10 \\

        $\pihigh$ & Probability that the unique optimal behavior pays off 1 
                & \textbf{0.9} \\

        $\pilow$ & Probability one of $B - 1$ non-optimal behaviors pays off 1 
                 & 0.1, 0.45, 0.8 \\

        $\tau$ & Softmax temp.; $\uparrow=$more exploration, $\downarrow=$more
                    exploitation 
               & 0.01, \textbf{0.1}, 1.0 \\

        $u$    & Probability optimal behavior changes between generations 
               & 0.0, 0.1, \ldots, 1.0 \\

        $L$    & Number of time steps per generation & 1, $B/2$, $B$, $2B$, $4B$ \\

        $N_T$    & Number of teachers to pool, from which best selected 
                 & \textbf{5}  \\

        \bottomrule
        \end{tabular} 
\end{table}

To guide behavior selection, agent $i$ counts how many times it
performed each behavior $b$, denoted $c_{ib}$. Agents use this count to 
update the mean payoff they have obtained from behavior $b$, denoted $\ledger$.
$\ledger$ is initialized to 0 for all $b$ at model start for
all agents (Figure~\ref{fig:schematic}A), and at 
generation start for asocial learners. Social
learning agents' $\ledger$ is initialized as that of its teacher, who is selected
through performance-biased oblique learning; $c_ib$ is set to 1 for all social
learners $i$ and behaviors $b$ (Figure~\ref{fig:schematic}C). Agent $i$'s
accumulated payoffs from performing several behaviors over their lifetime of $L$
time steps is denoted $\pi_{i}$. If an agent $i$ is a social learner, then we write
$s_i = 1$, otherwise the agent is an asocial learner and $s_i = 0$.

At generation start, one of the behaviors is set to 
yield expected payoff $\pihigh$, with the other $B-1$ behaviors yielding
expected payoffs $\pilow$. At each time step, agents perform behavior $b$ 
with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}},
\end{equation}
\noindent
where $\beta$ is a parameter that adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$)~\cite{McElreath2005}. 
Agents probabilistically recieve a payoff of 0 or 1 depending on whether the
bandit paid off, equivalent to a draw from a Bernoulli distribution with 
mean $\pi_b$. For concreteness we denote this probabilistic payoff
$\bandit$. On performing behavior $b$, agent $i$ updates the
corresponding behavior count by 1, $c_{ib} \leftarrow c_{ib} + 1$, and then
the expected payoffs calculated for that behavior are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}}.
\end{equation}
\noindent
This procedure continues for $L$ time steps within each generation
(Figure~\ref{fig:schematic}B).

In between generations agents first reproduce via asexual haploid reproduction; 
then new child agents learn from the
previous generation if they are social learners; and finally all agents from the
previous generation then die off (Figure~\ref{fig:schematic}C). 
$N$ reproducers are selected with replacement over $N$ independent draws, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
A child inherits its parent's social learning trait $s_i$ without mutation.
A social learner child with $s_i = 1$ learns from a teacher from its parent's
generation, including possibly their parent (oblique learning). 
A child selects its teacher by first selecting $N_T$ prospective
teachers from the population completely at random, then selecting the one with
highest net payoffs $\pi_i$ as their teacher, with ties broken randomly ($N_T =
5$ in the main text; \mt{Other values will be tested in the supplement}). While it
is somewhat arbitrary to first subset $N_T$, this is a conservative choice that
represnts the fact that access to the entire population is not generally guaranteed.
Social learners adopt their chosen teacher's ledger of expected payoffs $\ledger$,
and set all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$.  Asocial
learners ($s_i = 0$) are initialized with all $c_{ib} = 0$ and $\ledger = 0$ 
(Figure~\ref{fig:schematic}C). The
newly initialized population then begins behavior selection and payoff accumulation
for another $L$ time steps. The evolutionary process continues until $\sum_i s_i
= 1$ or $\sum_i s_i = 0$ \mt{I believe all trials will eventually end like this, so
in the next set of model runs I will not allow early termination before one of these
conditions are met}.

\begin{table}[h]
    \caption{Outcome variables.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.25in}p{0.85in}} \toprule

        Symbol & Description & Values \\

        \midrule

        $\meansl$ & Mean social learning prevalence over agents and trials
                  & $\in [0.0, 1.0]$ \\

        $\meanpi / L$ & Mean payoffs accumulated in a generation normalized by
        lifespan & $\in [0.0, 1.0]$ \\

        $\meanT / L$ & Mean number of generations to convergence & 20k / 8 max. \\
        \bottomrule
    \end{tabular}
\end{table}

To analyze the effect of the four principal uncertainty factors, we systematically
varied their values and observed three key outcome variables aggregated across 1000
trial simulations, and across all agents when appropriate. We varied $u \in \{0.0,
0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L
\in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We observed
three outcome measures: (1) $\meansl$, the average value of $s_i$ over all agents
and trials; (2) $\meanpi / L$, the average net payoffs at simulation end across
agents and trials, normalized by lifespan; and (3) $\meanT / L$, 
or the mean number of time steps to fixation normalized by
lifespan\footnote{Equivalent to the mean number of generations to fixation}
(fixation means $\sum_i s_i = 0 \text{ or } 1$). \mt{Make sure to qualitatively
introduce these and their evolutionary importance in the intro.}
We analyze outcomes by plotting $\meansl$ on the y-axis and environmental
variation is on x-axis since we theoretically expect that $\meansl$ will 
decrease monotonically from 1 to 0 as $u$ increases. We expected and observed
that the other three uncertainty factors shift the value of $u$ at which 
$\meansl$ starts to decrease, and how rapidly $\meansl$ decreases. Note that the
hypothesis-testing concept of significance is meaningless here because we could
make any small outcome difference ``significant'' by running more simulation trials.
We instead study the patterns of outcome 
variables over systematically varied uncertainty factor values. 
Our model was implemented in the Julia programming language~\cite{Bezanson2017} 
using the Agents.jl agent-based modeling library~\cite{Datseris2022} and run
on the Sherlock supercomputing cluster at Stanford University. Model code and
data are publicly available on GitHub\footnote{\url{https://github.com/mt-digital/UncMod}} 
and OSF.io\footnote{\mt{TODO}}.

@@

Several patterns emerged in changes to the shape of social learning suppression over
environmental variability, $u$, across different uncertainty contexts defined by the
payoff ambiguity, selection set size, and effective lifespan
(Figure~\ref{fig:mainResults}). When payoff ambiguity and selection set size were
both small ($\pilow = 0.1$ and $B=2$) social learning only evolved when $u$ is small
since individual learning is relatively efficient in this case
(Figure~\ref{fig:mainResults}, upper left). As $\pilow$ increased the suppression of
$\meansl$ flattened over $u$, which was more pronounced for smaller $L$ (rows of
Figure~\ref{fig:mainResults}).  When $L$ is smaller, agents have fewer opportunities
to learn through trial and error, and so it is more difficult for them to discern
which strategy, social or asocial, is optimal. Increased selection set size led to
both a shift in the critical value of $u$ when $\meansl$ begins to decrease, and
flattened social learning suppression. This is because possibly misleading 
social information is less consequential due
to (1) larger $B$ is a more difficult problem for trial-and-error learning and
(2) a misidentified optimal behavior is weighted less relative to smaller
$B$, meaning agents are more likely to explore behaviors identified as non-optimal
when $B$ is larger (columns of Figure~\ref{fig:mainResults}).  
Increased $B$ led to more pronounced flattening of social learning suppression when
$L$ was small, apparently again because of the limited data available to evolution,
making individual outcomes more path dependent.

@@

Social learning evolution was correlated with differences in population-level
expected payoffs for social versus asocial learning. Furthermore, we found that mean
payoffs realized by our simulated agents track either the expected social learning
payoffs, $\meansoc$, or the expected asocial learning payoffs, $\meanasoc$.
$\meansoc$ and $\meanasoc$ were calculated across uncertainty contexts by
initializing the model to be all social or all individual learners then running the
model for 100 generations for social learning and across 1000 one-generation trials
for individual learning.  Simulated evolutionary mean payoffs across agents and
trials, $\meanpi$, are initially higher than expected asocial learning, $\meanasoc$,
and comparable to expected social learning, $\meansoc$, in most cases 
(Figure~\ref{fig:payoffs}).  But as social learning becomes less beneficial due to
increased environmental variability, $\meansoc$ and $\meanpi$ approach $\meanasoc$.
It appears agents tend
to be conservative as expected social and asocial learning payoffs approach each
other---the transition to $\meansl = 0$ often happens although social
learning is still theoretically optimal. When $L=1$, $\meanpi$ can be greater than
$\meansoc$ because asocial learners at the beginning of the simulation help to 
pass on less misleading information on average, so that the cultural information passed on by
eventual all-social learner populations have more noise. This noise further reduces the
weighting of the previous generation's observed optimal payoff if it changes between
generations enabling agents to more efficiently explore alternatives before focusing
on a misidentified optimal behavior \mt{I think we could calculate this if it were
valuable}.

\begin{figure}
  \caption{Mean payoff normalized for lifespan (y-axes, solid lines with circles)
    generally track either expected all-social-learners payoff ($\meansoc$, dash-dot
    line with diamonds) or expected all-asocial payoff ($\meanasoc$, long dash
    horizontal lines) as $u$ increases (x-axes). When $\meansl$ goes from 1 to
  0 is when social learning payoffs approach individual learning payoffs due to
increased environmental variability.} 
  \label{fig:payoffs}
\centering
    \includegraphics[width=0.75\textwidth]{Figures/meanNetPayoffs.pdf}
\end{figure}

The number of generations to reach fixation increased
when expected social learning payoffs approached expected asocial payoffs, i.e.,
when $\meansl$ becomes suppressed (Figure~\ref{fig:steps}). Recall a fixated
state is when $\sum_i s_i = 0 \text{ or } 1$.  This supports our explanation that
the supression of social learning begins when individual simulation outcomes are
more path dependent and susceptible to end in a non-optimal state since simulations
in these conditions also take longer to reach fixation.  For example, when
$\pilow=0.1$, $B=4$, and $L=8$, we see a sharp increase in $\meanT$ when social learning
suppression begins; but when $L=1$, social learning suppression is flatter
than for $L=8$, and  correspondingly its $\meanT$ is not as strongly peaked, but is
elevated across the $u$ (Figure~\ref{fig:steps} upper left).  In some cases $L=1$
trials reached fixation faster than $L=8$ or $L=20$; larger $L$ had larger peak $\meanT/L$
than smaller $L$, for example, both for the cases shown in Figure~\ref{fig:steps}
and across all settings tested \mt{TODO: add full figure to SI}. 
However, we also see situations where $L=1$ and fixation takes longer than $L=8$.
This is especially pronounced for
$\pilow = 0.45$ where social learning suppression is much flatter for $L=1$
(Figure~\ref{fig:steps}, bottom row). In general, then, a sharper peak in $\meanT$
indicates a sharper suppression of social learning over $u$ and a flatter peak,
elevated over more values of $u$, indicates greater path dependence and more 
drift overall.

@@

By identifying four common uncertainty dimensions we developed a model that predicts
how different forms of uncertainty might interact to foster or suppress social
learning.  The model reproduced classic predictions that social learning can yield
higher payoffs than individual learning when uncertainty is sufficiently limited.
By disambiguating and organizing various operationalizations of uncertainty, 
we developed a more nuanced theoretical understanding of which forms of
uncertainty interact in which ways to make social learning more or less beneficial
compared to asocial learning.  Note that in many cases we made modeling choices that
excluded some important mechanisms of social learning; for example we assumed
success-biased learning, although conformity is theoretically important as
well~\cite{Muthukrishna2016a,Smaldino2018b}. Still, we believe our model
advances theoretical parsimony; and the software implementation is designed to
facilitate testing alternate modeling choices.

\paragraph{Disc. 2: Foundation for studying social learning evolution with group
structure or other topics(PAUL, JAMIE, CRISTINA)}~\cite{Katsnelson2014}.

Our modeling approach may have applications to developing social artificial
intelligence. Much effort has gone in to developing the cognitive machinery to
realize multi-agent reinforcement learning
systems~\cite{Sandholm1996,Ndousse2021,Gronauer2022}, with some work explicitly
incorporating cultural evolutionary approaches to social learning~\cite{Jaques2019}.
However, it seems there are a lack of methods for adapting to environmental
uncertainty, or determining when social learning is actually advantageous under
different uncertainty conditions. Our model dynamics could be used as an
evolutionary algorithm to update AI agents' social learning strategy without
explicit knowledge of, or need to approximate, the underlying uncertainty structure.
Our model could also predict the expected benefit of social learning, which may be
inherently costly in artificial multi-agent learning systems, and how long the
translated evolutionary algorithm may take to converge.

\bibliographystyle{apacite}
% \bibliography{/Users/mt/workspace/Writing/library.bib}
\bibliography{this.bib}

\appendix

\section{Appendix}

\subsection{Convergence information}
